---
title: "netflix_hmw"
author: "EmreCanSi"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: yes
    toc_depth: '2'
    df_print: paged
    code_folding: hide
  html_notebook:
    toc: yes
    toc_depth: 2
    toc_float: yes
---


```{r, include = FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(lubridate)
```

## READING DATA
```{r}
netflixraw <- read_csv("https://github.com/ygterl/EDA-Netflix-2020-in-R/raw/master/netflix_titles.csv")

```


## Understanding DATA
Then we can review the data with following codes:glimpse and summary

```{r}
library(dplyr)
glimpse(netflixraw)

```


```{r}
summary(netflixraw)
```
```{r}
netflix <- netflixraw %>% select(show_id,type,title,director,cast,rating,country,release_year,rating,duration,listed_in)
```



```{r}
netflix <- distinct(netflix, title, country, type, release_year, .keep_all = TRUE)

summary(netflix)
```

in here we seperate values with comma and accept them as different raws.


```{r}
netflixallcountry<-netflix %>% 
    mutate(country = strsplit(as.character(country), ", ")) %>% 
    unnest(country)

netflixallcountry1<-netflixallcountry %>% group_by(country) %>% summarise(country_number = n())

netflixallcountry2<-netflixallcountry1[order(netflixallcountry1$country_number),,decreasing=TRUE]

netflixalllisted_in<-netflix %>% 
    mutate(listed_in = strsplit(as.character(listed_in), ", ")) %>% 
    unnest(listed_in)

netflixalllisted_in1<-netflixalllisted_in %>% group_by(listed_in) %>% summarise(listed_in_number = n())

netflixalllisted_in2<-netflixalllisted_in1[order(netflixalllisted_in1$listed_in),,decreasing=TRUE]

netflixalllisted_in1 <- filter(netflixalllisted_in2, listed_in_number>200)
```

Most created type of listed_in is below:

```{r}
library(dplyr)
library(ggplot2)


  ggplot(netflixalllisted_in1)+geom_point(aes(x = netflixalllisted_in1$listed_in, y = netflixalllisted_in1$listed_in_number, color=as.character(listed_in)))


```
While international movies is top of the list, dramas, comedies and international TV shows are follows the international TV shows.

As you can see below there are more tv show in Netflix in 2019

```{r}

moviesandtvshows <- netflixraw %>%
  filter(release_year >= 1980 & release_year < 2020) %>%
  group_by(type, release_year) %>%
  summarise(count = n()) %>%
  arrange(desc(release_year)) %>%
  ungroup()
moviesandtvshows

  ggplot(moviesandtvshows)+geom_line(aes(x = moviesandtvshows$release_year, y = moviesandtvshows$count, color=as.character(type)))

```

Here is the top 20 artist that love to play in Netflix.

```{r}


netflixartists<-netflix %>% 
    mutate(cast = strsplit(as.character(cast), ", ")) %>% 
    unnest(cast)

netflixartists1<-netflixartists %>% group_by(cast) %>% summarise(numberofplay = n())
 
netflixartists1 %>% top_n(20, numberofplay)

```
